title: "Covid91 vaccine study Final2023"
author: "Jared Lobitana"
output:
html_document:
number_sections: true
toc: true
toc_depth: 5
date: "`r Sys.Date()`"
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new preventative vaccine (covid921) is under test in clinical trials.
Given the nature of this vaccine it is suspected that the drug may exhibit
different effectiveness across the variety of subpopulations tracked.
Consider population subgroups by gender( guys and gals), drug use, and
sexual orientation. Write a complete report that separately considers this
drug for each subgroup and your recommendations as to whether people in the
subgroup should take the vaccine. Structure your report in four sections
each devoted to one of the four subgroups, use null and alternative
hypothesis statements and correct statistical verbage when referencing
p-values in conclusions.

data from this study is taken from FinalData


$H_0$: the vaccines has no effect on people within a subgroup

$H_A$: the vaccine has an effect on people within a subgroup


## Methods

The statistical analysis i will be using to sort the data will be cat~cat
because we are putting the sub groups into category



the methods we will use to display the data is bar chart, cross table, and
chi squared test

## sub categorize male data
```{r}
maleData = filter (FinalData,sex == "M")
```


## Males


## Graphical Results
```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")

```

This graph shows use that from the all male group of people who contracted
the ones who received a vaccine had less symptoms then those with the
placebo while the all male group that didnt contract covid showed no
difference between those who were vaccinated and un vaccinated.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=maleData)
rowPerc(table1)
colPerc(table1)
```

The data shows that 69.57% of the non vaccinated male group experienced
symptoms while only 30.43% of the vaccinated males experienced symptoms.
It also tells us that from the male placebo group 2.55% of them contracted
COVID whilst only 1.15% of the vaccinated group contracted COVID.

## Inferential Result


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p value for the chi squared test is .0000000000001098 which indicates quite a significant relationship between the variables therefore disproving the null hypothesis $H_A$ giving strong evidence for the presence of an association between the variables. For the fisher exact test the pvalue is .00000000000004631 giving a strong basis for the observed association between the variables to unlikely have occurred by chance alone. It also gives us a Odds ratio showing theres a 2.242522% higher chance of the events odds happening in one group giving more evidence to the positive association of the variables and rejection of the null $H_A$



## sub categorizing female data

```{r}
femaleData = filter (FinalData,sex == "F")
```

## Females

## Graphical Results

```{r}
barchartGC(~infected + treatment, data=femaleData)
barchartGC(~infected+treatment, data=femaleData, type="percent")

```

This graph shows use that from the all female group of people who had contracted COVID-19 the ones who received a vaccine had less symptoms overall then those with the placebo. while the all female group that didn't contract COVID showed almost no difference between those who were vaccinated and un vaccinated.

## Numerical Results

```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```

The data from this table shows that from the female group that were infected with covid-19 of those who were unvaxed 60% of them experienced symptoms while only 40% of people in the vaccinated group experienced symptoms.

Furthermore, this data also depicts that within the female placebo group 2.04% were infected with covid but only 1.37% of the vaccinated group.

## Inferential Result

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The data suggests a significant relationship between the treatment type (placebo vs. vaccine) and infection status based on Pearson's Chi-squared test (p-value = 0.0003). Therefore, we reject the null hypothesis $H_A$ , indicating that there is an association between the variables.

Fisher's exact test for count data on table2 reveals a odds ratio of (1.493076) with a 95% confidence interval of 1.199509 to 1.862988 suggesting a statistically significant positive association between the two categorical variables, indicating that the event's odds are about 1.49 times higher in one group compared to the other.

## LGBTQ



## Druggies


# Overall Results and Conclusions