title: "Covid91 vaccine study Final2023"
author: "Jared Lobitana"
output:
html_document:
number_sections: true
toc: true
toc_depth: 5
date: "`r Sys.Date()`"
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new preventative vaccine (covid921) is under test in clinical trials.
Given the nature of this vaccine it is suspected that the drug may exhibit
different effectiveness across the variety of subpopulations tracked.
Consider population subgroups by gender( guys and gals), drug use, and
sexual orientation. Write a complete report that separately considers this
drug for each subgroup and your recommendations as to whether people in the
subgroup should take the vaccine. Structure your report in four sections
each devoted to one of the four subgroups, use null and alternative
hypothesis statements and correct statistical verbage when referencing
p-values in conclusions.

data from this study is taken from FinalData


$H_0$: the vaccines has no effect on people within a subgroup

$H_A$: the vaccine has an effect on people within a subgroup


## Methods

The statistical analysis i will be using to sort the data will be cat~cat
because we are putting the sub groups into category



the methods we will use to display the data is bar chart, cross table, and
chi squared test



## sub categorize male data
```{r}
maleData = filter (FinalData,sex == "M")
```


## Males


## Graphical Results
```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")

```

This graph shows use that from the all male group of people who contracted
the ones who received a vaccine had less symptoms then those with the
placebo while the all male group that didnt contract covid showed no
difference between those who were vaccinated and un vaccinated.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=maleData)
rowPerc(table1)
colPerc(table1)
```

The data shows that 69.57% of the non vaccinated male group experienced
symptoms while only 30.43% of the vaccinated males experienced symptoms.
It also tells us that from the male placebo group 2.55% of them contracted
COVID whilst only 1.15% of the vaccinated group contracted COVID.

## Inferential Result


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p value for the chi squared test is .0000000000001098 which indicates quite a significant relationship between the variables therefore disproving the null hypothesis $H_A$ giving strong evidence for the presence of an association between the variables. For the fisher exact test the pvalue is .00000000000004631 giving a strong basis for the observed association between the variables to unlikely have occurred by chance alone. It also gives us a Odds ratio showing theres a 2.242522% higher chance of the events odds happening in one group giving more evidence to the positive association of the variables and rejection of the null $H_A$



## sub categorizing female data

```{r}
femaleData = filter (FinalData,sex == "F")
```

## Females

## Graphical Results

```{r}
barchartGC(~infected + treatment, data=femaleData)
barchartGC(~infected+treatment, data=femaleData, type="percent")

```

This graph shows use that from the all female group of people who had contracted COVID-19 the ones who received a vaccine had less symptoms overall then those with the placebo. while the all female group that didn't contract COVID showed almost no difference between those who were vaccinated and un vaccinated.

## Numerical Results

```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```

The data from this table shows that from the female group that were infected with covid-19 of those who were unvaxed 60% of them experienced symptoms while only 40% of people in the vaccinated group experienced symptoms.

Furthermore, this data also depicts that within the female placebo group 2.04% were infected with covid but only 1.37% of the vaccinated group.

## Inferential Result

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The data suggests a significant relationship between the treatment type (placebo vs. vaccine) and infection status based on Pearson's Chi-squared test (p-value = 0.0003). Therefore, we reject the null hypothesis $H_A$ , indicating that there is an association between the variables.

Fisher's exact test for count data on table2 reveals a odds ratio of (1.493076) with a 95% confidence interval of 1.199509 to 1.862988 suggesting a statistically significant positive association between the two categorical variables, indicating that the event's odds are about 1.49 times higher in one group compared to the other.

The p-value from fishers exact test (0.0002162) indicates strong evidence against the null hypothesis, suggesting a statistically significant association between the two categorical variables again disproving the null $H_A$ 

## subsetting lgtbq data

```{r}
LGBTQData = filter (FinalData,LGBTQ == "gay")
```
## LGBTQ
## Graphical Results

```{r}
barchartGC(~infected+treatment, data=LGBTQData)
barchartGC(~infected+treatment, data=LGBTQData, type="percent")

```

This Graph shows that within the LGBTQ subgroup Those afflicted with covid 19 and that were vaccinated experienced more symptoms then their un vaccinated counter part. for the people who never contracted COVID, vaxxed or unvaxxed, there was no difference shown

## Numerical Results

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```

This table shows that within the all gay placebo group 15.38% experienced symptoms while 84.62% of the vaccinated group had symptoms

From the all gay placebo group only .66% of the unvaccinated got infected with COVID-19 while a large 3.54% of vaccinated gay people got infected with COVID-19.

## Inferential Result
```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The p value within the LGBTQ group is 0.0000006668 showing that theres a significant association between the variables so i am rejecting the null hyothesis $H_A$ and conclude the covid vaccine has an effect within the LGBTQ population

The Fishers exact tesst p value is 1.029e-08, which shows evidence for the extremely low probability of obtaining the observed results by chance alone, suggesting a highly significant association between the variables therefore disproving the null hypothesis $H_A$

The odds ratio of the fisher exact test is 0.181787 indicating that the event's odds are approximately 0.18 times lower in one variable group compared to the other, implying a significant difference between the two groups. Again adding to the body of evidence against the null hyopthesis $H_A$

## subset druggie data
```{r}
DrugUserData = filter(FinalData, DrugUser == "yes")
```
## Druggies
## Graphical Results

```{r}
barchartGC(~infected + treatment, data=DrugUserData)
barchartGC(~infected+treatment, data=DrugUserData, type="percent")

```

This graph shows that for druggies who were infected with COVID-19 those who were vaccinated actually experienced more symptoms then those that were un-vaccinated. The druggies who didn't get infected with COVID-19 whether they were vaccinated or un-vaccinated, it showed an insignificant difference.

## Numerical Results

```{r}
table4 <- xtabs(~infected + treatment, data=DrugUserData)
rowPerc(table4)
colPerc(table4)
```

The table shows that within the placebo druggie group, 11.67% of these people experienced COVID-19 symptoms when they were infected with COVID-19, whilst 88.33% of vaccinated drug users experienced symptoms.

The bottom table shows that only .4% of the placebo group got infected with COVID, whilst 3.06% of the people from the vaccinated group got infected with COVID.

## Inferential Result

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The p-value of .000000002842 from the Pearson's Chi-squared test shows us a highly significant association between the variables, which concludes with the rejection of the null hypothesis $H_A$.

Fisher's exact test yielded a highly significant p-value (2.32e-10) and a 95% confidence interval (0.0482 to 0.2791), which indicates a powerful connection with the categorical variables, and with the estimated odds ratio being 0.126, this implies that the likelihood of this event happening is significantly lower in one group compared to the other, which is supported by a 95% confidence interval, ranging from 0.04819608 to 0.27914768, and based on this amount of evidence, I am lead to the conclusion of rejecting the null hypothesis $H_A$ .


# Overall Results and Conclusions
In conclusion, the effectiveness of the COVID vaccine varied between the different subgroups. Staring with the all male group, we are rejecting the null hypothesis. I came to this conclusion because based on the bar chart, numerical graph, and the chisq and fisher tests, these all showed that the men who received the COVID vaccine and contracted it ended up experiencing less symptoms than those who received the placebo. The men that did not get vaccinated and did not contract COVID showed little to no difference between each other.As for the group that received it, the placebo had an infection rate of 2.55%, and those of them who got the vaccine had an infection rate of 1.15%, showing that the vaccine is effective in men, which concludes that we are rejecting the null hypothesis for the male group.

Moving onto the all female group, we are also rejecting the null hypothesis. I came to this conclusion because based on on all of the tests that had been run, they showed that the women who received the COVID vaccine and contracted it also experienced less symptoms than the women who received the placebo. The female group that did not get vaccinated and did not contract COVID also showed little to no difference compared to each other. Within the group that did receive the placebo, they had an infection rate of 2.04%. The group that received the vaccine had an infection rate of 1.37%, concluding that the vaccine is effective in women, therefore rejecting the null hypothesis for the female subgroup.

When the male and female subgroups are compared, women had a smaller percentage of infection with the COVID vaccine and placebo than then men by 0.22% and 0.51%. While this finding isn't a big difference, it still shows that within the data analysis women experience less symptoms with the COVID vaccine.

In the LGBTQ subgroup, we are rejecting the null hypothesis. Based on our tests, we can conclude that the COVID vaccine is effective in this subgroup, but negatively. We found out that the LGBTQ people who received the COVID vaccine and contracted it experienced significant symptoms than those who received the placebo. The LGBTQ group that did not get vaccinated and did not contract COVID were almost exactly the same, showing no difference. In the LGBTQ group that did receive the placebo had an infection rate of 0.66% and those who did receive the vaccine had an infection rate of 3.54%, a very significant difference between the two, therefore rejecting the null hypothesis since the vaccine had an effect on the group.

Finally, for the drug user group, they received the same result as the LGBTQ subgroup. We are rejecting the null hypothesis due to the drug users experiencing significantly more symptoms than the placebo group, showing that the COVID vaccine did have an effect, but not a positive one In the group of drug users that were vaccinated and contracted COVID, they had a 3.06% infection rate. The placebo group had a 0.4% infection rate showing a very big difference between the two and how much symptoms each group got, once again, we are rejecting the null.

In conclusion, when comparing all of the tests from most positive COVID affects to the more negative COVID affects, the females ended up having the best results with the COVID vaccine, while the LGBTQ subgroup ended up with the worst results, where they experienced the most symptoms whilst they had COVID, even after having taken the vaccine. Males had almost the same results as the females, just a little less and drug users had fewer symptoms by only 0.48%. So based on these findings, I'd suggest for cis males and females, to get the COVID vaccine seeing as the positive effect outweighed the negative effects, but for the drug users and members of the LGBTQ ccommunity, I would suggest not taking it as the opposite effects were found for them with the negative effects highly outweighing the positive effects. 