title: "Covid91 vaccine study Final2023"
author: "Jared Lobitana"
output:
html_document:
number_sections: true
toc: true
toc_depth: 5
date: "`r Sys.Date()`"
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new preventative vaccine (covid921) is under test in clinical trials.
Given the nature of this vaccine it is suspected that the drug may exhibit
different effectiveness across the variety of subpopulations tracked.
Consider population subgroups by gender( guys and gals), drug use, and
sexual orientation. Write a complete report that separately considers this
drug for each subgroup and your recommendations as to whether people in the
subgroup should take the vaccine. Structure your report in four sections
each devoted to one of the four subgroups, use null and alternative
hypothesis statements and correct statistical verbage when referencing
p-values in conclusions.

data from this study is taken from FinalData


$H_0$: the vaccines has no effect on people within a subgroup

$H_A$: the vaccine has an effect on people within a subgroup


## Methods

The statistical analysis i will be using to sort the data will be cat~cat
because we are putting the sub groups into category



the methods we will use to display the data is bar chart, cross table, and
chi squared test

## sub categorize male data
```{r}
maleData = filter (FinalData,sex == "M")
```


## Males


## Graphical Results
```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")

```

This graph shows use that from the all male group of people who contracted
the ones who received a vaccine had less symptoms then those with the
placebo while the all male group that didnt contract covid showed no
difference between those who were vaccinated and un vaccinated.


## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions